FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-11 NewVersion/paper.tex
title
Added a small example. Now S3 is ready.

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,updates},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\ttfamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

@@

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

>>>>

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,override},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\ttfamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

@@

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{FHJ: A Formal Model for Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

### 2017-10-11 NewVersion/sections/Formalization.tex
Added a small example. Now S3 is ready.

\subsection{Subtyping and Typing Rules}
\paragraph{Subtyping}
The subtyping of \MIM{} consists of only a few rules shown at the top of Figure~\ref{fig:typingrules}.
In short, subtyping relations are built from the inheritance in interface
declarations. Subtyping is both symmetric, reflexive and transitive.

@@

\textsc{(T-Invk)} is the typing rule for method invocation.
Naturally, the receiver and the arguments are required to be well-typed.
$\mbody$ is our key function for method lookup, based on the
hierarchical dispatching algorithm. The formal definition will be introduced in
Section~\ref{subsec:auxdefs}. To overview,
the hierarchical invocation $\mbody(m, I, J)$ means ``finding most specific $m$
above $I$ and along path $J$'', and usually $I$ denotes the dynamic type while $J$
denotes the static type. In \textsc{(T-Invk)}, as the compilation should not be aware
of the dynamic type, it only requires that invoking $m$ is valid for the static type of the
receiver. The result of $\mbody$ contains the interface that provides the most specific implementation,
the parameters and the return type. We use underscore for the return expression, to imply that an empty return expression
from an abstract method is also acceptable.
Note that we accept implicit upcasts on the arguments.
% \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

\textsc{(T-New)} is the typing rule for object creation $\new{I}$. The auxiliary function $\canInstantiate(I)$ checks whether interface $I$ can be instantiated or not, formalized in Section~\ref{subsec:auxdefs}. Since triangle inheritance accepts conflicting branches to coexist, the check requires that the most specific method is concrete for each method on each branch.

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function from Section~\ref{subsec:auxdefs},
used to find ``the most specific interfaces that are above $I$ and along path $J$, and originally defines $m$''. By ``most specific interfaces'',
it implies that the inherited super types are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates the characteristics of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
needs to be type-checked, and there is also an implicit cast to the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In the condition we again use some helper functions from Section~\ref{subsec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any super type $J$ and any method $m$.

@@

\paragraph{Examples for mbody: TODO}

\begin{figure*}
\saveSpaceFig

@@

>>>>

\subsection{Subtyping and Typing Rules}\label{subsec:typingrules}
\paragraph{Subtyping}
The subtyping of \MIM{} consists of only a few rules shown at the top of Figure~\ref{fig:typingrules}.
In short, subtyping relations are built from the inheritance in interface
declarations. Subtyping is both symmetric, reflexive and transitive.

@@

\textsc{(T-Invk)} is the typing rule for method invocation.
Naturally, the receiver and the arguments are required to be well-typed.
$\mbody$ is our key function for method lookup, based on the
hierarchical dispatching algorithm. The formal definition will be introduced in
Section~\ref{sec:auxdefs}. To overview,
the hierarchical invocation $\mbody(m, I, J)$ means ``finding most specific $m$
above $I$ and along path $J$'', and usually $I$ denotes the dynamic type while $J$
denotes the static type. In \textsc{(T-Invk)}, as the compilation should not be aware
of the dynamic type, it only requires that invoking $m$ is valid for the static type of the
receiver. The result of $\mbody$ contains the interface that provides the most specific implementation,
the parameters and the return type. We use underscore for the return expression, to imply that an empty return expression
from an abstract method is also acceptable.
Note that we accept implicit upcasts on the arguments.
% \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

\textsc{(T-New)} is the typing rule for object creation $\new{I}$. The auxiliary function $\canInstantiate(I)$ checks whether interface $I$ can be instantiated or not, formalized in Section~\ref{sec:auxdefs}. Since triangle inheritance accepts conflicting branches to coexist, the check requires that the most specific method is concrete for each method on each branch.

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function from Section~\ref{sec:auxdefs},
used to find ``the most specific interfaces that are above $I$ and along path $J$, and originally defines $m$''. By ``most specific interfaces'',
it implies that the inherited super types are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates the characteristics of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
needs to be type-checked, and there is also an implicit cast to the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In the condition we again use some helper functions from Section~\ref{sec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any super type $J$ and any method $m$.

@@

\paragraph{Counter-Example} Counter-examples can promote understanding on the typing rules. Here we refer to case (e)
in Figure~\ref{fig:examplesmbody}. To be specific, the entire program in \MIM{} is as follows:

\vspace{3pt}\begin{lstlisting}
interface T                 {}
interface T1 extends T      {}
interface T2 extends T      {}
interface A                 { T m() override A { return new T(); } }
interface B extends A       { T m() override A { return new T1(); } }
interface C extends A       { T m() override A { return new T2(); } }
interface D extends B, C {}

((A) new D()).m()
\end{lstlisting}\vspace{3pt}
This program does not compile on $D$, because of the second condition in \textsc{(T-Intf)}, where $I$ equals $D$ and $J$ equals $A$.
By the algorithm, $\mbody(m, A, A)$ will refer to $A.m$, but $\mbody(m, D, A)$ is undefined, since both $B.m$ and $C.m$ are most specific
to D along path $A$, which forms a diamond. The expression \lstinline|((A) new D()).m()| is one example of triggering ambiguity, but \MIM{}
simply rejects the definition of $D$. To resolve the issue, the programmer need to have an overriding method in $D$, to explicitly merge
the conflicting ones.

\begin{figure*}[t]
\saveSpaceFig

@@

\begin{comment}
\paragraph{Example} In contrast with the counter-example in Section~\ref{subsec:typingrules}, it is better to understand semantics by
well-compiled examples. Here we abstract a variant of the \lstinline|DrawableDeck| example:

\vspace{3pt}\begin{lstlisting}
interface Void       {}
interface JFrame     {}
interface Deck       { Void draw() override Deck { return new Void(); } }
interface Drawable { JFrame draw() override Drawable; }
interface DrawableDeck extends Drawable, Deck {
  JFrame draw() override Drawable {
    return new JFrame();
  }
}

((Drawable) new DrawableDeck()).draw()
\end{lstlisting}\vspace{3pt}
We put \lstinline|Drawable.draw| as an abstract method instead, but hierarchically override it in \lstinline|DrawableDeck|.
By typing rules, the code is well-compiled. And during runtime,
\begin{align*}
	& ((Drawable) new DrawableDeck()).draw() \\
\rightarrow & (JFrame) new JFrame()
\end{align*}
\end{comment}

### 2017-10-11 NewVersion/sections/Overview.tex
added a UML graph
Added a small example. Now S3 is ready.

\subsection{Problem 1: Unintentional Method Conflicts}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| method in \lstinline|Deck| simply
prints out a message.

@@

\begin{figure*}[t]
	\saveSpaceFig
	\centering
	\includegraphics[height=2cm]{pics/DrawableSafeDeck.pdf}
	\caption{UML graph for \lstinline|DrawableSafeDeck|.}\label{fig:drawablesafedeck}
	\saveSpaceFig
\end{figure*}

>>>>

\subsection{Problem 1: Unintentional Method Conflicts}\label{subsec:problem1}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| method in \lstinline|Deck| simply
prints out a message.

@@

% \begin{figure*}[t]
% 	\saveSpaceFig
% 	\centering
% 	\includegraphics[height=2cm]{pics/DrawableSafeDeck.pdf}
% 	\caption{UML graph for \lstinline|DrawableSafeDeck|.}\label{fig:drawablesafedeck}
% 	\saveSpaceFig
% \end{figure*}

\begin{figure*}[t]
  % \nocaptionrule
  \centering
  \begin{minipage}[t]{0.4\textwidth}
  \includegraphics[height=4cm]{pics/DrawableSafeDeck1.pdf}
  \end{minipage}
  \hspace*{2pt}
  \begin{minipage}[t]{0.5\textwidth}
  \includegraphics[height=4cm]{pics/DrawableSafeDeck0.pdf}
  \end{minipage}
  \caption{UML graphs for \lstinline|DrawableSafeDeck|.}\label{fig:drawablesafedeck}
\end{figure*}
\yanlin{todo: change text and add a ref in Introduction}
